---
title: "PCA_LiMM"
author: "Ren Xie"
date: "2/19/2021"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(warning = FALSE)

```
load in data
```{r 1, echo = T, eval = T}
library(ggplot2)
library(lme4)
library(reshape2)
library(dplyr)
library(factoextra)
################import data ##################################
load("D:/Bioinformatics and system biology/YEAR 2/Age group/Project/Linear mixed model/dts_hotel_project/load_to_R/data.Rdata")
common_columns = colnames(data$data_r)[1:3] #select columns of categorical variables
chylo.colnames <-  colnames(data$data_r)[grep('XL_VLDL_.{0,3}.$', colnames(data$data_r))] #select metabolites of chylomicron XXL AND XL VLDL

chylo.original= data$data_r[c(common_columns, chylo.colnames)]

#Missing value checking
sapply(chylo.original, function(x) sum(is.na(x))) #check for missing value
chylo.original[rowSums(is.na(chylo.original)) > 0,]#show the row with missing value
```
Standardization and preprocessing

```{r 2, echo = T, eval = T, warning= FALSE}
################standardization by individual variance from baseline (time point 1) #############
baseline <- filter(chylo.original, chylo.original$Time == 1) #select value at time point 1
baseline.sd <- apply(baseline[,-c(1:3)], 2, sd)#calculate standard deviation of baseline

#standardize the data
chylo.standardized <- chylo.original
chylo.standardized[,-c(1:3)] <- sweep(chylo.standardized[,-c(1:3)],2,baseline.sd,'/') 
chylo <- chylo.standardized 

################preprocess the data #############
chylo$treatment <- stringr::str_split_fixed(chylo$ID, '_', 2)[,2] #create treatment column
chylo$treatment = as.factor(chylo$treatment)

car_var <- c('measurements', 'particpant', 'treatment', 'ID', 'num_timing', 'Time', 'PCs') #list of categorical variables in dataset
#visulize the dataset
fig_tl = ggplot(melt(chylo, id.vars = c('particpant', 'treatment' , 'Time'), measure.vars = chylo.colnames, variable.name = 'measurements'), aes(x = Time, y = value , group = interaction(particpant, treatment) )) +
  facet_wrap(~measurements) +
  geom_line(aes(colour = treatment))
fig_tl

```
PCA the standardized raw data (without missing value)
```{r 3} 

chylo.pca <- prcomp(chylo[!colnames(chylo) %in% car_var], scale. = TRUE) #perform PCA
#Suggestion:using robust PCA (robust SVD based) dataframe with missing value (pcaMethods from bioconductor is a good package for that)

#scree plot
fviz_eig(chylo.pca, title = 'PCA for raw data') 

#pc explained
chylo.PCexplain <- data.frame('value'= as.vector(chylo.pca$sdev^2/sum(chylo.pca$sdev^2)), 
                                          row.names = colnames(chylo.pca$rotation)) #store the explanation rates per PC
```

Plot the score plot
```{r 4}
n_PCs.perserved = 3
n_PCs.perserved.names <- paste0('PC',seq(1:n_PCs.perserved))
chylo.score <- as.data.frame(chylo.pca$x[,n_PCs.perserved.names])
chylo.score <- cbind(chylo[colnames(chylo) %in% car_var],chylo.score )
chylo.score.long <- melt(chylo.score, id.vars = c('particpant', 'treatment' , 'Time'), measure.vars = n_PCs.perserved.names,variable.name = 'PCs' )
chylo.score.long$particpant <- as.factor(chylo.score.long$particpant)
chylo.score.long$treatment <- as.factor(chylo.score.long$treatment)
chylo.score.long$PCs <- as.factor(chylo.score.long$PCs)

pc = 'PC1'
chylo.score.perPC <- dplyr::filter(chylo.score.long, PCs == pc)
fig <- ggplot(chylo.score.perPC, aes(x = Time, y = value, group = treatment))
fig <- fig + geom_point(aes(x = Time, y = value, colour = treatment))
fig <- fig + scale_color_manual('Treatment', values = c('A' = 'red', 'B' = 'blue'))
fig <- fig + scale_x_continuous(breaks=seq(1,11))
fig <- fig + xlab('Time point') + ylab(paste0(pc, ' (', round(chylo.PCexplain[pc,] * 100, 2), '%)'))

fig

pc = 'PC2'
chylo.score.perPC <- dplyr::filter(chylo.score.long, PCs == pc)
fig <- ggplot(chylo.score.perPC, aes(x = Time, y = value, group = treatment))
fig <- fig + geom_point(aes(x = Time, y = value, colour = treatment))
fig <- fig + scale_color_manual('Treatment', values = c('A' = 'red', 'B' = 'blue'))
fig <- fig + scale_x_continuous(breaks=seq(1,11))
fig <- fig + xlab('Time point') + ylab(paste0(pc, ' (', round(chylo.PCexplain[pc,] * 100, 2), '%)'))

fig

pc = 'PC3'
chylo.score.perPC <- dplyr::filter(chylo.score.long, PCs == pc)
fig <- ggplot(chylo.score.perPC, aes(x = Time, y = value, group = treatment))
fig <- fig + geom_point(aes(x = Time, y = value, colour = treatment))
fig <- fig + scale_color_manual('Treatment', values = c('A' = 'red', 'B' = 'blue'))
fig <- fig + scale_x_continuous(breaks=seq(1,11))
fig <- fig + xlab('Time point') + ylab(paste0(pc, ' (', round(chylo.PCexplain[pc,] * 100, 2), '%)'))

fig



```
Loading plot
```{r}
loading.map <- factoextra::fviz_pca_var(chylo.pca,geom = c('point', 'text') , repel = TRUE, gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),col.var = 'black') + 
    xlab(paste0('PC1 (', round(chylo.PCexplain['PC1',] * 100, 2), '%)')) + ylab(paste0('PC2 (', round(chylo.PCexplain['PC2',] * 100, 2), '%)')) + theme_bw() + ggtitle('Measurement loading')
loading.map
```

create frame for desgin matrix
```{r 3, echo = T, eval = T, results='asis' }

#############3 degree model desgin matrix ##############
#loading polynomial functions
source('../../Software/DifferentFunctionalForms.R')

#making dummy dataframe of time points
dummies_data = data.frame(diag(11)) 
dummies_colnames = ''
for (i in 1:11) {
  dummies_colnames[i] = paste0("time",i)
}
colnames(dummies_data) = dummies_colnames
dummies_data$num_timing=seq(1,11)

#set knot and make dummies for desgin matrix
knot1 = 2
dummies = linear_third_pw_myway(dummies_data, knot1)
knitr::kable(dummies, "pipe")
```

Model building
``````{r 5, echo = T, eval = T, warning = TRUE}
################model building#############
data.melt <- chylo.score.long

#create empty list for parameters
fix_parameters.list <- list()
rand_parameters <- list()

#create dataframe for  Akaike information criterion (AIC), evaluation of model likelihood 
df.aic <-  data.frame(row.names = unique(data.melt$PCs))
df.aic$AIC <- 0
lm_list <- list()

#build models per metabolite in the dataset
for (pc in  levels(data.melt$PCs)) {

  cur_position <- match(pc, unique(data.melt$PCs)) #current position in the loop
  print(paste0(cur_position, "/",length(unique(data.melt$PCs)))) #showing the progress
  temp_pc = dplyr::filter(data.melt, PCs == pc) #create temporary subset of a metabolite
  colnames(temp_pc)[which(colnames(temp_pc) == 'Time')] = 'num_timing'  #since num_timming is default colnames in piece wise linear model
  temp_pc = plyr::join(temp_pc, dummies) #joining temporary dataset with dummies
  
  lm_temp = lmer(value ~  num_timing1*treatment  + num_timing2*treatment + num_timing1b*treatment + num_timing3 * treatment +  treatment+
                   (num_timing1*treatment + num_timing2*treatment +  num_timing1b*treatment + num_timing3 * treatment +  treatment |particpant ),
                 data = temp_pc,  REML=FALSE)
  lm_list[[pc]] <- lm_temp #save the model
  rand_parameters[[pc]] <-  ranef(lm_temp) #save the parameters of random effect
  df.aic[pc,] <-  AIC(logLik(lm_temp)) #save the AIC value
  fix_parameters.list[[pc]] <-  fixef(lm_temp) #save the parameters of fixed effect
}

```
Extracted fixed effect
```{r 5, echo = T, eval = T, results='asis' }
###################################extract fix effect##################
df.fix_parameters <- data.frame() #create data frame for random effect of interaction between treatment and individual
for(pc in names(fix_parameters.list)){
  #show the progress
  # cur_position <- match(mea, names(fix_parameters.list)) #find your current position
  # print(paste0(cur_position, "/",length(names(fix_parameters.list)))) #show the progress
  
  #unpack the list of random effect parameters
  pc.fix_parameters <- as.data.frame(t(fix_parameters.list[[pc]]))  #extract 'treatment and individual interaction' random effect parameter per PCs
  
  #add additional information
  pc.fix_parameters$PCs <- as.factor(pc) #add PCs column
  
  
  df.fix_parameters <- rbind(df.fix_parameters,pc.fix_parameters)#add individual dataframe into whole data frame 'interaction '
}
row.names(df.fix_parameters) <- df.fix_parameters$PCs #change the rownames

df.fix_parameters <- as.data.frame(t(df.fix_parameters))

#calculate fixed effect matrix
fix_effect <- data.frame()
fix_effect.percolumn <- data.frame()
for (pc in names(lm_list)){
  temp.fix_effect <- model.matrix(lm_list[[pc]]) %*% fixef(lm_list[[pc]]) #fixed effect
  temp.fix_effect.percolumn <- sweep(model.matrix(lm_list[[pc]]), 2, fixef(lm_list[[pc]]), '*') #fixed effect per column of design matrix
  
  temp.fix_effect <- data.frame('Estimated_fixed'=temp.fix_effect)
  temp.fix_effect.percolumn <- as.data.frame(temp.fix_effect.percolumn)
  
  
  temp.fix_effect$PCs <- as.factor(pc)
  temp.fix_effect.percolumn$measurements <- as.factor(pc)
  
  
  temp.fix_effect <- cbind(dplyr::filter(data.melt, PCs == pc)[1:3], temp.fix_effect)#adding categorical variable
  temp.fix_effect.percolumn <- cbind(dplyr::filter(data.melt, PCs == pc)[1:3], temp.fix_effect.percolumn)#adding categorical variable
  
  fix_effect <- rbind(fix_effect, temp.fix_effect)
  fix_effect.percolumn <- rbind(fix_effect.percolumn, temp.fix_effect.percolumn)
}
fix_effect <- dplyr::filter(fix_effect, particpant == 'P001') #fixed effect is same for every subjects, so only one of them is enough
summary(fix_effect)
```
visualization fixed effect
``````{r 6, echo = T, eval = T, results='asis' }
###################################Visualization fixed effect #################
#visualize the fixed effect

fig <- ggplot(fix_effect)
fig <- fig + geom_line(aes(x = Time, y = Estimated_fixed, colour = treatment))
fig <- fig + facet_wrap(~PCs)
fig <- fig + scale_color_manual('Treatment', values = c('A' = 'red', 'B' = 'blue'))
fig <- fig + scale_x_continuous(breaks = unique(fix_effect$Time))
fig <- fig + xlab('Time point') + ylab('Standardized concentration')
fig

```
Extract random effect
```{r 7, echo = T, eval = T, results='asis' }
#########################Extract random effect###########
random_parameters.list <- rand_parameters #should be a list
random_effect.percolumn <- data.frame()
random_effect <- data.frame()
for (pc in names(lm_list)){
  temp_matrix <- model.matrix(lm_list[[pc]]) #extract design matrix
  temp_matrix <- cbind(dplyr::filter(data.melt, PCs == pc)[colnames(data.melt) %in% car_var[!car_var %in% 'PCs']], temp_matrix)#combined with categorical variable
  temp_random_parameter <-  as.matrix(random_parameters.list[[pc]]$particpant) #extract dataframe of random parameters of one metabolite
  #extract random parameters for each subject
  for (indi in unique(data.melt$particpant)){
    temp_matrix.indi <- dplyr::filter(temp_matrix, particpant == indi)[!colnames(temp_matrix) %in% car_var] #design matrix for single subject
    temp.random_effect.percolumn <- sweep(as.matrix(temp_matrix.indi), 2, temp_random_parameter[indi,], '*')#column wise random effect (random effect per feature)
    temp.random_effect <- as.matrix(temp_matrix.indi) %*% temp_random_parameter[indi,]#calculate total random effect
    
    temp.random_effect <- data.frame('Estimated_random'=temp.random_effect) #create dataframe for store random effect
    temp.random_effect.percolumn <- as.data.frame(temp.random_effect.percolumn)
    
    temp.random_effect.percolumn$PCs <- as.factor(pc)
    temp.random_effect$PCs <- as.factor(pc)
    
    temp.categorical <- dplyr::filter(temp_matrix, particpant == indi)[colnames(temp_matrix) %in% car_var[!car_var %in% 'PCs']]
    temp.random_effect.percolumn <- cbind(temp.categorical, temp.random_effect.percolumn)
    temp.random_effect <- cbind(temp.categorical, temp.random_effect)
    
    random_effect.percolumn <- rbind(random_effect.percolumn, temp.random_effect.percolumn)
    random_effect <- rbind(random_effect, temp.random_effect)
  }
}
```
Visualization random effect
```{r 8, echo = T, eval = T, results='asis' }
##################Visualization random effect #################
#construct a matrix with random effect per individual
fig <- ggplot(random_effect)
fig <- fig + geom_line(aes(x = Time, y = Estimated_random, colour = treatment,linetype = PCs))
fig <- fig + facet_wrap(~particpant)
fig <- fig + scale_color_manual('Treatment', values = c('A' = 'red', 'B' = 'blue'))
fig <- fig + scale_x_continuous(breaks = unique(random_effect$Time))
fig <- fig + xlab('Time point') + ylab('Value')
fig

```
Significance test (exclude one individual from score (not from raw data and re-PCA))
```{r 9, echo = T, eval = T, results='asis', warning=F }
################Significance test ##########
Model_quality.list <- list() #Quality of models
sig.rand_parameters.list <- list() #list of lists of parameters of random effect
sig.fix_parameters.list <- list()# list of lists of  parameters of fixed effect
sig.lm_list <- list() #list of lists of models

for (indi in levels(data.melt$particpant)){ # loop to exclude one subject out of dataset
  
  cur_position <- match(indi, levels(data.melt$particpant)) #position 
  print(paste0(indi, ": " ,cur_position, "/",nlevels(data.melt$particpant))) #print the postion of progress
    
  temp.melt <- dplyr::filter(data.melt, !particpant == indi) #One subject excluded from the dataset

  #create temporary a list and a dataframe for storing parameters
  temp.sig.fix_parameters.list<- list()
  temp.sig.rand_parameters.list <- list()
  temp.model_quality <-  data.frame(row.names = unique(data.melt$PCs))
  temp.model_quality$AIC <- 0
  temp.model_quality$Loglik <- 0
  temp.lm_list <- list()
  
  
  for (pc in  levels(data.melt$PCs)) { #loop through all metabolites
    cur_position <- match(pc, unique(data.melt$PCs)) #position 
    print(paste0(pc, ": " ,cur_position, "/",length(unique(data.melt$PCs)))) #print the postion of progress
    
    sig.temp_mea = dplyr::filter(temp.melt, PCs == pc) #extract temporary dataset for one measuremnt
    colnames(sig.temp_mea)[which(colnames(sig.temp_mea) == 'Time')] = 'num_timing'  #since num_timming is default colnames in piece wise linear model
    sig.temp_mea = plyr::join(sig.temp_mea, dummies) #join with dummies
    
    sig.lm_temp = lmer(value ~  num_timing1*treatment  + num_timing2*treatment + num_timing1b*treatment + num_timing3 * treatment +  treatment+
                     (num_timing1*treatment + num_timing2*treatment +  num_timing1b*treatment + num_timing3 * treatment +  treatment |particpant ),
                   data = sig.temp_mea,  REML=FALSE) # build model for each subset
    temp.lm_list[[pc]] <- sig.lm_temp #store the modle
    temp.sig.rand_parameters.list[[pc]] <-  ranef(sig.lm_temp) #store random effect
    temp.sig.fix_parameters.list[[pc]] <-  fixef(sig.lm_temp) #store fixed effect

    temp.model_quality[pc,'AIC'] <-  AIC(logLik(sig.lm_temp)) #store AIC score
    temp.model_quality[pc,'Loglik'] <-  logLik(sig.lm_temp) #store log Likelyhood 
  }
  Model_quality.list[[indi]] <- temp.model_quality 
  sig.rand_parameters.list[[indi]] <- temp.sig.rand_parameters.list #store list of random parameter from exclude one subject into a total list
  sig.fix_parameters.list[[indi]] <- temp.sig.fix_parameters.list #store list of fixed  parameter from exclude one subject into a total list
  sig.lm_list[[indi]] <- temp.lm_list
  
}
```

significance test: extract fixed effect 
```{r 10}
################significance test: extract fixed effect ##########
sig.fix_effect.percolumn <- data.frame() #set a data frame to store fixed effect (per feature of design matrix) from significant test 
sig.fix_effect <- data.frame() #set a data frame to store fixed effect from significant test 


for (indi in names(sig.lm_list)){
  gather.fix_effect.percolumn <- data.frame() #a temporary dataset for per feature 
  gather.fix_effect <- data.frame() 

  
  for (pc in names(sig.lm_list[[indi]])){
    temp.sig.fixed_effect.percolumn <- sweep(model.matrix(sig.lm_list[[indi]][[pc]]), 2, fixef(sig.lm_list[[indi]][[pc]]), '*') #column wise multiple with the parameter, calculate fixed effect per feature without add them together
    temp.sig.fixed_effect <- model.matrix(sig.lm_list[[indi]][[pc]]) %*% fixef(sig.lm_list[[indi]][[pc]]) #calculate the fixed effect
    
    temp.sig.fixed_effect.percolumn <- data.frame('value' = temp.sig.fixed_effect.percolumn) 
    temp.sig.fixed_effect.percolumn <- cbind(dplyr::filter(data.melt, PCs == pc & (!particpant == indi))[1:4], temp.sig.fixed_effect.percolumn)
    temp.sig.fixed_effect.percolumn$one_leave <- as.factor(indi)
    gather.fix_effect.percolumn <- rbind(gather.fix_effect.percolumn, temp.sig.fixed_effect.percolumn)
    
    temp.sig.fixed_effect <- data.frame('value'=temp.sig.fixed_effect)
    temp.sig.fixed_effect <- cbind(dplyr::filter(data.melt, PCs == pc & (!particpant == indi))[1:4], temp.sig.fixed_effect)
    temp.sig.fixed_effect$one_leave <- as.factor(indi)
    gather.fix_effect <- rbind(gather.fix_effect, temp.sig.fixed_effect)
    
  }
  #since fixed effect is same for all subject, so only one of them will be kept
  if ('P001' %in% unique(gather.fix_effect.percolumn$particpant)) {
    gather.fix_effect.percolumn <- dplyr::filter(gather.fix_effect.percolumn, particpant == 'P001')
    sig.fix_effect.percolumn <- rbind(sig.fix_effect.percolumn, gather.fix_effect.percolumn)
    
    gather.fix_effect <- dplyr::filter(gather.fix_effect, particpant == 'P001')
    sig.fix_effect <- rbind(sig.fix_effect, gather.fix_effect)
    
  } else{
    gather.fix_effect.percolumn <- dplyr::filter(gather.fix_effect.percolumn, particpant == 'P002')
    sig.fix_effect.percolumn <- rbind(sig.fix_effect.percolumn, gather.fix_effect.percolumn)
    
    gather.fix_effect <- dplyr::filter(gather.fix_effect, particpant == 'P002')
    sig.fix_effect <- rbind(sig.fix_effect, gather.fix_effect)
  }
}

summary(sig.fix_effect)

```
significance test: extract random effect
```{r 11}
#####################significance test: extract random effect###################
sig.random_effect.percolumn <- data.frame() #set a data frame to store random effect (per feature of design matrix) from significant test 
sig.random_effect <- data.frame() #set a data frame to store random effect from significant test 

#progress bar
pb <- progress::progress_bar$new(
  format = "  Progress [:bar] :percent",
  total = length(sig.rand_parameters.list), clear = FALSE, width= 60)

#extract random effect
for (indi_left in names(sig.rand_parameters.list)){
  pb$tick()
  Sys.sleep(1 / 100)
  
  # cur_position <- match(indi_left, names(sig.rand_parameters.list)) #position 
  # print(paste0(cur_position, "/",length(sig.rand_parameters.list))) #print the potion of progress
  
  #loop through all the methods
  for (pc in names(sig.lm_list[[indi_left]])){
    temp.matrix <- model.matrix(sig.lm_list[[indi_left]][[pc]]) #extract design matrix
    temp.matrix <- cbind(dplyr::filter(data.melt, PCs == pc & (!particpant == indi_left))[1:4], temp.matrix) #extract the part of design matrix of the certain measurement and without the left out subject
    temp.sig.random_parameter <-  as.matrix(sig.rand_parameters.list[[indi_left]][[pc]]$particpant) #extract the corresponding parameters
    
    for (indi in unique(temp.matrix$particpant)){
      temp.matrix.indi <- dplyr::filter(temp.matrix, particpant == indi)[!colnames(temp.matrix) %in% car_var] #extract the part of design matrix for certain individual
      temp.sig.random_effect.percolumn <- sweep(as.matrix(temp.matrix.indi), 2, temp.sig.random_parameter[indi,], '*') #calculate random effect per feature
      temp.sig.random_effect <- as.matrix(temp.matrix.indi) %*% temp.sig.random_parameter[indi,] #calculate total random effect
      
      temp.sig.random_effect.percolumn <- data.frame('value'= temp.sig.random_effect.percolumn)
      temp.sig.random_effect.percolumn <- cbind(dplyr::filter(temp.matrix, particpant == indi)[1:4], temp.sig.random_effect.percolumn)
      temp.sig.random_effect.percolumn$one_leave <- as.factor(indi_left)
      sig.random_effect.percolumn <- rbind(sig.random_effect.percolumn, temp.sig.random_effect.percolumn)
      
      temp.sig.random_effect <- data.frame('value'= temp.sig.random_effect)
      temp.sig.random_effect <- cbind(dplyr::filter(temp.matrix, particpant == indi)[1:4], temp.sig.random_effect)
      temp.sig.random_effect$one_leave <- as.factor(indi_left)
      sig.random_effect <- rbind(sig.random_effect, temp.sig.random_effect)
    }
  }
}

summary(sig.random_effect)

```
Calculate interval for fixed effect
```{r 12}
score.intervals <- data.frame()
for(pc in n_PCs.perserved.names){
  temp.score <- dplyr::filter(sig.fix_effect, PCs == pc)
  temp.fixed_effect.score <- dplyr::filter(fix_effect, PCs == pc)
  
  temp.score$particpant <- NULL
  temp.fixed_effect.score$particpant <- NULL
  temp.score <- tidyr::spread(temp.score, one_leave, value)
  temp.score <- plyr::join(temp.score, temp.fixed_effect.score)
  
  temp.intervals <- temp.score[colnames(temp.score) %in% car_var]
  temp.intervals$PCs <- as.factor(pc)
  temp.intervals$q2.5 <- apply(temp.score[!colnames(temp.score) %in% car_var], 1, function(x) quantile(x,.025, na.rm = TRUE))
  temp.intervals$q97.5 <- apply(temp.score[!colnames(temp.score) %in% car_var], 1, function(x) quantile(x,.975,na.rm = TRUE))
  temp.intervals$sd <- apply(temp.score[!colnames(temp.score) %in% car_var], 1, function(x) sd(x, na.rm = TRUE))
  
  score.intervals <- rbind(score.intervals, temp.intervals)
}
```

Visualization fixed effect with error bar
```{r 13}
#########################Visualization score of fixed effect##############
fixed_effect.withinterval <- plyr::join(fix_effect, score.intervals)

pc = 'PC1'
fixed_effect.score.perPC <- dplyr::filter(fixed_effect.withinterval, PCs == pc)
fig <- ggplot(fixed_effect.score.perPC, aes(x = Time, y = Estimated_fixed, group = treatment))
fig <- fig + geom_point(aes(x = Time, y = Estimated_fixed, colour = treatment))
fig <- fig + geom_errorbar(aes(ymin=q2.5, ymax= q97.5,colour = treatment ), width=.1)
fig <- fig + scale_color_manual('Treatment', values = c('A' = 'red', 'B' = 'blue'))
fig <- fig + scale_x_continuous(breaks=seq(1,11))
fig <- fig + xlab('Time point') + ylab(paste0(pc, ' (', round(chylo.PCexplain[pc,] * 100, 2), '%)'))

fig

pc = 'PC2'
fixed_effect.score.perPC <- dplyr::filter(fixed_effect.withinterval, PCs == pc)
fig <- ggplot(fixed_effect.score.perPC, aes(x = Time, y = Estimated_fixed, group = treatment))
fig <- fig + geom_point(aes(x = Time, y = Estimated_fixed, colour = treatment))
fig <- fig + geom_errorbar(aes(ymin=q2.5, ymax= q97.5,colour = treatment ), width=.1)
fig <- fig + scale_color_manual('Treatment', values = c('A' = 'red', 'B' = 'blue'))
fig <- fig + scale_x_continuous(breaks=seq(1,11))
fig <- fig + xlab('Time point') + ylab(paste0(pc, ' (', round(chylo.PCexplain[pc,] * 100, 2), '%)'))

fig

pc = 'PC3'
fixed_effect.score.perPC <- dplyr::filter(fixed_effect.withinterval, PCs == pc)
fig <- ggplot(fixed_effect.score.perPC, aes(x = Time, y = Estimated_fixed, group = treatment))
fig <- fig + geom_point(aes(x = Time, y = Estimated_fixed, colour = treatment))
fig <- fig + geom_errorbar(aes(ymin=q2.5, ymax= q97.5,colour = treatment ), width=.1)
fig <- fig + scale_color_manual('Treatment', values = c('A' = 'red', 'B' = 'blue'))
fig <- fig + scale_x_continuous(breaks=seq(1,11))
fig <- fig + xlab('Time point') + ylab(paste0(pc, ' (', round(chylo.PCexplain[pc,] * 100, 2), '%)'))

fig


```
